%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{Interactive QTL charts within R Markdown}

## Interactive QTL charts within R Markdown

[R/qtlcharts](http://kbroman.github.io/qtlcharts) is an R package to
create interactive charts for xQTL data, for use with
[R/qtl](http://www.rqtl.org).
While its interactive QTL charts are particularly useful for
exploratory data analysis within R, they may also be included within 
[R Markdown](https://www.rstudio.com/ide/docs/r_markdown)-based reports.

[R Markdown](https://www.rstudio.com/ide/docs/r_markdown) is a
extension of [Markdown](http://daringfireball.net/projects/markdown/),
using [knitr](http://yihui.name/knitr/), to write simply-marked-up
text with R code chunks that are converted to an html file, for
viewing in the browser. Code chunks are processed by knitr and
replaced with the results or graphs produced. 

For more information on R Markdown and knitr, see the
[knitr in a knutshell](http://kbroman.github.io/knitr_knutshell/)
tutorial.

To include [R/qtlcharts](http://kbroman.github.io/qtlcharts)-based
interactive charts within an R Markdown document, you include the
relevant calls to the R function within a code chunk, using the chunk
option `results="asis"`. Also, include `print=TRUE` and likely
`onefile=TRUE` in the function call. Below, I'll illustrate how it's
done.

### An example

As an example, we'll consider the `grav` dataset included with
R/qtlcharts. These are data from
[Moore et al. Genetics 195:1077-1086, 2013](http://www.ncbi.nlm.nih.gov/pubmed/23979570),
on a QTL experiment on gravitropism in Arabidopsis, with data on 162
recombinant inbred lines (Ler &times; Cvi). Seedlings were sprouted
and then rotated 90 degrees with respect to gravity; the growth of the
seedlings was then recorded on video. The outcome is the root tip
angle (in degrees) at two-minute increments over eight hours.

We'll first load the R/qtlcharts package and the data set.

```{r load_data, message=FALSE}
library(qtlcharts)
data(grav)
```

For this illustration, we'll reduce the size of the data set, by
retaining only every 5th phenotype value.

```{r subset_pheno}
grav$pheno <- grav$pheno[,seq(1, nphe(grav), by=5)]
```

The times at which the measurements were made are recorded (in
minutes) in the phenotype names. We'll grab those times and convert
them to hours.

```{r grab_times}
times <- as.numeric(sub("T", "", phenames(grav)))/60
```

Let's further run a single-QTL genome scan with each individual time
point, using Haley-Knott regression.

```{r scanone}
grav <- calc.genoprob(grav, step=1)
out.hk <- scanone(grav, pheno.col=1:nphe(grav), method="hk") 
```

### Interactive charts

To use the interactive charts from R/qtlcharts within an R Markdown
document, you must first load all of the javascript-based
resources. (When the charts are created from within R, this is done
separately, and automatically, with each function call. But within an
R Markdown document, we want to do it just once.)

This is done by calling `print_qtlcharts_resources()` within a code
chunk, using the chunk option `results="asis"`. I would generally
suppress the code using `echo=FALSE`, but here we'll show the
details.

```{r print_qtlcharts_resources, results="asis"}
print_qtlcharts_resources()
```

We're now free to begin making charts. With each, you'll want to use
`print=TRUE`. And you **don't** want to do that within R. You may be
sitting watching a megabyte of text fly (or crawl) by.

In addition, you want to use the argument `chartdivid` to define some
unique character string for each interactive plot. These are sort of
like the names of code chunks with knitr.

Finally, each of these code chunks is with `results="asis"`.


### Now, the charts


First, let's make an interactive chart of the genetic map of markers
for these data, just because it's easy. `iplotMap` will take either a
map object or the cross object `grav`; in the latter case, it uses
the R/qtl function `pull.map` to pull out the map.

```{r iplotMap, results="asis"}
iplotMap(grav, print=TRUE)
```

Now, let's use `iplotCorr` to plot a heat map of the correlation
matrix for the phenotype at all pairs of times, linked to scatterplots
of the phenotypes.

```{r iplotCorr, results="asis"}
iplotCorr(grav$pheno, title="", chartdivid="iplotCorr",
          print=TRUE)
```

Next, let's use `plotCurves` to create an interactive plot of the
phenotype traces, linked to scatterplots of two cross-sections.
We use `chartOpts` to specify the axis labels.

```{r iplotCurves, results="asis"}
iplotCurves(grav$pheno, times,
            grav$pheno[,c("T30", "T240")],
            grav$pheno[,c("T240", "T480")],
            chartdivid="iplotCurves",
            print=TRUE,
            chartOpts=list(curves_xlab="Time", curves_ylab="Root tip angle",
              scat1_xlab="Angle at 30 min", scat1_ylab="Angle at 4 hrs",
              scat2_xlab="Angle at 4 hrs", scat2_ylab="Angle at 8 hrs"))
```


Finally, `iplotMScanone` produces a heat map of the LOD scores for
each individual time point at each genomic position, linked to the
individual LOD curves and to a plot of the estimated QTL effect over
time.

```{r iplotMScanone, results="asis"}
iplotMScanone(out.hk, grav, chartdivid="iplotMScanone",
              print=TRUE, chartOpts=list(eff_ylab="QTL effect"))
```

### A bit of CSS

One final detail: I'm including a bit of
[CSS](http://en.wikipedia.org/wiki/Cascading_Style_Sheets) code to
revise the left margin for the document that's produced. This is done
as straight text, not inside a code chunk.

```
<style type="text/css">
body { margin-left: 100px; }
</style>
```

<style type="text/css">
body { margin-left: 100px; }
</style>
