---
title: R/qtlcharts within R Markdown
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Interactive QTL charts within R Markdown}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8](inputenc)
---

[R/qtlcharts](http://kbroman.org/qtlcharts) is an R package to
create interactive charts for QTL data, for use with
[R/qtl](http://www.rqtl.org).
While its interactive QTL charts are particularly useful for
exploratory data analysis within R, they may also be included within
[R Markdown](http://rmarkdown.rstudio.com)-based reports.

[R Markdown](http://rmarkdown.rstudio.com) is a
extension of [Markdown](http://daringfireball.net/projects/markdown/),
using [knitr](http://yihui.name/knitr/), to write simply-marked-up
text, with chunks of R code, that is converted to an html file, for
viewing in a web browser. The chunks of code are processed by knitr and
replaced with the results or graphs produced.

For more information on R Markdown and knitr, see the
[knitr in a knutshell](http://kbroman.org/knitr_knutshell/)
tutorial.

To include [R/qtlcharts](http://kbroman.org/qtlcharts)-based
interactive charts within an R Markdown document, you just need to include the
relevant calls to the R function within a code chunk.

I'm going to use defaults of `fig.width=8` and `fig.height=6` for
each of these charts.

```{r knitr_options}
knitr::opts_chunk$set(fig.width=8, fig.height=6)
```


### An example

As an example, we'll consider the `grav` dataset included with
R/qtlcharts. These are data from
[Moore et al. Genetics 195:1077-1086, 2013](http://www.ncbi.nlm.nih.gov/pubmed/23979570),
on a QTL experiment on gravitropism in Arabidopsis, with data on 162
recombinant inbred lines (Ler &times; Cvi). Seedlings were sprouted
and then rotated 90 degrees with respect to gravity; the growth of the
seedlings was then recorded on video. The outcome is the root tip
angle (in degrees) at two-minute increments over eight hours.

We'll first load the R/qtl and R/qtlcharts packages and the data set.

```{r load_data, message=FALSE}
library(qtl)
library(qtlcharts)
data(grav)
```

For this illustration, we'll reduce the size of the data set, by
retaining only every 5th phenotype value.

```{r subset_pheno}
grav$pheno <- grav$pheno[,seq(1, nphe(grav), by=5)]
```

The times at which the measurements were made are recorded (in
minutes) in the phenotype names. We'll grab those times and convert
them to hours.

```{r grab_times}
times <- as.numeric(sub("T", "", phenames(grav)))/60
```

Let's further run a single-QTL genome scan with each individual time
point, using Haley-Knott regression.

```{r scanone}
grav <- calc.genoprob(grav, step=1)
out.hk <- scanone(grav, pheno.col=1:nphe(grav), method="hk")
```



### Now, the charts

First, let's make an interactive chart of the genetic map of markers
for these data, just because it's easy. `iplotMap` will take either a
map object or the cross object `grav`; in the latter case, it uses
the R/qtl function `pull.map` to pull out the map.

```{r iplotMap}
iplotMap(grav)
```

Now, let's use `iplotCorr` to plot a heat map of the correlation
matrix for the phenotype at all pairs of times, linked to scatterplots
of the phenotypes. (Note that for this plot, I'm using
`fig.height=4.2` in order to avoid some empty space below the figure.)

```{r iplotCorr, fig.height=4.2}
iplotCorr(grav$pheno)
```

Next, let's use `plotCurves` to create an interactive plot of the
phenotype traces, linked to scatterplots of two cross-sections.
We use `chartOpts` to specify the axis labels.

```{r iplotCurves}
iplotCurves(grav$pheno, times,
            grav$pheno[,c("T30", "T240")],
            grav$pheno[,c("T240", "T480")],
            chartOpts=list(curves_xlab="Time", curves_ylab="Root tip angle",
                           scat1_xlab="Angle at 30 min", scat1_ylab="Angle at 4 hrs",
                           scat2_xlab="Angle at 4 hrs", scat2_ylab="Angle at 8 hrs"))
```


Finally, `iplotMScanone` produces a heat map of the LOD scores for
each individual time point at each genomic position, linked to the
individual LOD curves and to a plot of the estimated QTL effect over
time.

```{r iplotMScanone}
iplotMScanone(out.hk, grav, chartOpts=list(eff_ylab="QTL effect"))
```

### Charts within a loop

If you want to produce multiple interactive charts within a loop, you
need to finesse it a bit with `tags` and `tagList` from the
[htmltools](https://github.com/rstudio/htmltools) package.

Create a list containing the charts, wrap each within a `<p>` tag
using `tags$p()`, and then use `tagList` when you print the list.

```{r iplotScanone_loop}
times <- c("T30", "T240", "T480")
times_number <- match(times, phenames(grav))
plot_list <- vector("list", length(times))
for(i in seq(along=times))
    plot_list[[i]] <- iplotScanone(out.hk, lodcolumn=times_number[i])
plot_list <- lapply(plot_list, htmltools::tags$p)
htmltools::tagList(plot_list)
```
